Storm Surges

### 2014-07-18 RoughWork/manuscript.tex
Included description of model winds for feb 2006

\subsection{Factors Contributing to Storm Surges: Feb 4, 2006}
%Include spatial extend of surge

@@

Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented in Figure \ref{fig:factors}. The modelled residual from three simulations have been plotted. First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly, is shown in blue. Additionally, a simulation with all forcing except the atmospheric forcing is presented in green and a simulation with all forcing except the sea surface height anomaly is shown in red. It is clear that the sea surface height anomaly contributes most significantly to the surge at each of these locations. When the sea surface height anomaly is not included as a forcing condition, the surge drops from 70 cm to 15.6 cm at Point Atkinson. Additionally, the timing of the maximum surge occurs much later. A similar drop in the maximum surge and delay in the timing of the maximum surge is observed at the other locations. Note that the press

The contribution to the surge due to atmospheric forcing appears to be geographically dependent, its removal leading to a drop of about 5 cm in the maximum surge at Point Atkinson and 10 cm at Campbell River. When atmospheric forces are neglected the timing of the maximum surge is earlier at Point Atkinson but later at Campbell River. Campbell River exhibits a strong response to atmopsheric conditions during midday on Feb 4, as demonstrated in a two sharp spikes in both the blue and red curves arounf 12:00 hr The green curve, which does not include any atmopsheric focring, does not exhibit this beahvious. The result is a peak surge earlier than in day than the no atmospheric forcing case. This response is also noted in the observations at Campbell River in Figure \ref{fig:feb2006} where the maxmium surge occurs earlier than at the other locations. At Victoria and Patricia Bay, the surge amplitude is mainly unaffected by the atmospheric forcing.  
%maybe look at winds/pressure to see why Vic/Pat Bay aren't affected?. Plan: map of wind vectors from CGRF (not for paper, just for me in aiding with explanation)

@@

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/dec2006_spatial.pdf}
\caption{Propagation of the storm surge on Dec 14/15 2006. Modelled residual every 2 hours starting on Dec 14 at 18:30. Time advances from top left to bottom right. Note that the images are aligned with the numerical grid which is rotated $29^{\circ}$ counter-clockwise of North. }
\label{fig:spatial}
\end{figure}

>>>>

\subsection{Factors Contributing to Storm Surges: Feb 4, 2006}

@@

Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented in Figure \ref{fig:factors}. The modelled residual from three simulations have been plotted. First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly, is shown in blue. Additionally, a simulation with all forcing except the atmospheric forcing is presented in green and a simulation with all forcing except the sea surface height anomaly is shown in red. It is clear that the sea surface height anomaly contributes most significantly to the surge at each of these locations. When the sea surface height anomaly is not included as a forcing condition, the surge drops from 70 cm to 15.6 cm at Point Atkinson and its maximum occurs eight hours later.  A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations. The atmospheric model indicates southeasterly winds during the onset of the surge but after the peak surge has passed, the winds shift to westerly. A strong westerly wind would cause an elevated sea surface height at Point Atkinson as water piles against the coast, explaining the delay in peak surge when only atmospheric forcing is used since the winds to shift westerly after the peak surge has passed. A simulation with only wind forcing (no inverse barometer effect or anomaly at open boundary) produces a surge of 9.0 cm at Point Atkinson, a small fraction of the total observed surge. However, this wind-induced surge occurs well after the peak surge from observations and simulations with full forcing.

The contribution to the surge due to atmospheric forcing appears to be geographically dependent, its removal leading to a drop of about 5 cm in the maximum surge at Point Atkinson and 10 cm at Campbell River. When atmospheric forces are neglected the timing of the maximum surge is earlier at Point Atkinson but later at Campbell River. Campbell River exhibits a strong response to atmopsheric conditions during midday on Feb 4, as demonstrated in a two sharp spikes in both the blue and red curves arounf 12:00 hr The green curve, which does not include any atmopsheric focring, does not exhibit this behaviour. The result is a peak surge earlier than in day than the no atmospheric forcing case. This response is also noted in the observations at Campbell River in Figure \ref{fig:feb2006} where the maxmium surge occurs earlier than at the other locations. At Victoria and Patricia Bay, the surge amplitude is mainly unaffected by the atmospheric forcing. 
%maybe look at winds/pressure to see why Vic/Pat Bay aren't affected?. Plan: map of wind vectors from CGRF (not for paper, just for me in aiding with explanation)

@@

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/dec2006_spatial.pdf}
\caption{Propagation of the storm surge on Dec 14/15 2006. Modelled residual every 2 hours starting on Dec 14 at 18:30. Time advances from top left to bottom right. Note that the images are oriented with the numerical grid which is rotated $29^{\circ}$ counter-clockwise of North. }
\label{fig:spatial}
\end{figure}
